Towards a classification framework for social machines

### 2013-02-26 latex/paper_new.tex
Fix quotation marks all being end-quotes

This paper provides a framework for...

@@

\section{Social machines: an introduction}
Once upon a time 'machines' were programmed by programmers and used by users. The success of the Web has changed this relationship: we now see configurations of people interacting with content and with each other, typified by social Web sites. Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them  and  treat  each  such  compound  as  a  'social  machine'  -  a  machine in which the two aspects are seamlessly interwoven. This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.

Essentially social machines can be characterized as assemblies of manually executed and machine-driven (as in 'automatized') services and the interaction of such services. A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship. When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines  because the users  are  part  of  the  'social  computation', in this case the data creation and collection process facilitate through the site.  This social component  becomes  richer when  the database  is  curated by members  of  the broader community (e.g., Wikipedia) and when  the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through their individual or joint activities.

@@

This paper makes two main contributions: First, we introduce the notion of 'social machines' and position it in connection to similar ideas and initiatives. Second we propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.

The remaining sections are organized as follows. In order to pin down what makes up a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} gives a summary of this comparative analysis. In Section \ref{sec:framework} we then present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via grid analysis. We conclude with an outlook of our future development activities, including details about the evaluation of the framework. In addition, we plan to engage with the broader community to set up a 'social machine' bringing together researchers from various disciplines who will contribute to the 'computation' of a joint and widely agreed classification framework for the field.

\section{Comparison with related fields}
\label{sec:comparison}
Since the invention of the Turing machine we have experienced a 'paradigm shift' in the usage of computers, moving away from purely calculative devices to facilitators of a wide range of human interactions. The emergence of 'Computer-supported Collaboration Work (CSCW)' \cite{grudin1994computer} is representative for the early days of this trend. CSCW addresses 'how collaborative activities and their coordination can be supported by means of computer systems' \cite{grudin1994computer}, whereas the initial concept evolved towards the more broader field of 'Computer-supported collaboration' (CSC).

The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web. These developments led to amazing growth in terms of the amounts of content available online and mass participation. They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and lively taking part in defining the agenda of many aspects of our society. This trend towards 'prosumerism' is finding more and more adopters in the public and private sector as well, as governments and enterprises not only become active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organizational management, product development, services offers, and policies. In this context, a number of terms are used to refer to the ways people interact with each other and with applications: 'wisdom of the crowds', 'collective intelligence', 'open innovation', 'crowdsourcing', 'human computation', and 'social computing'. They are related, but not synonymous with 'social machines', as explained in the following (see also Figure \ref{fig:socialmachine}).

@@

Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales hardly conceivable in the past, and to involve highly diverse and geographically distributed participants. A similar concept, though broader scoped, is collective intelligence, defined in  \cite{malone2009harnessing} as 'groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as 'social machines' is hidden in the the automation part of the latter. In a social machine, human and computational intelligence coalesce in order to achieve a given purpose. The overall setting may slightly vary in terms of the interaction patterns between the two types of services and the way their outcomes are used; however, where wisdom of the crowds and collective intelligence clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realization of hybrid systems where the two types of components co-exist. As such, theoretical and empirical insights from the two related areas are useful to understand the dynamics of the social structures underlying such a system, but they are definitely not the only ingredient needed to build operational social machines. The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering exercise by which a system is developed, tested, and updated are equally important.

One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a 'a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}. At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, while in the same time ensuring in-time access of a potentially infinite pool of skills and resources not available before the advent of social computing technologies. The term crowdsourcing is typically associated with this larger collection of situations, in which 'a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}. Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding. This sort of tasks are an important part of today's crowdsourcing projects landscape, in particular on so-called 'microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work in the range of seconds to minutes to complete.  The key difference between social machines and open innovation and crowdsourcing is again in the computational part. The novelty these two approaches bring in lies in using a much larger pool of human resources as traditional work environments \cite{quinn2011human}. There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution. A principled combination of human and computational capabilities is less in the focus of such systems and the technical means that support them, by contrast to social machines. In comparison to human computation, social machines cover a wider range of scenarios. Human computation is AI-centric and uses people to perform tasks that computers are not (yet) able to tackle (in terms of accuracy) \cite{quinn2011human}; by comparison, we see many successful examples of social machines in which the role of machines is rather to facilitate interactions within groups of people or communities of interest.

An analogous line of reasoning could be followed to point out the overlaps between social machines and social computing \cite{parameswaran2007research}. The latter is an area of computer science which refers systems that support 'the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets' \cite{parameswaran2007research}. As such, compared to the general concept of 'Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort. This distinction is even stronger in the case of social machines, which look at the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.

@@

\subsection{Methodology}
\label{sec:methodology}
The classification framework presented in this paper was created through knowledge elicitation ~\cite{knowledgeelicitation}. In particular we used the {\it repertory grid} elicitation
technique~\cite{kelly} in order to derive an initial set of {\it elements}, which represent instances of social machines, and {\it constructs}, which capture their most important characteristics.
In repertory grid elicitation a software tool is used to ask users to describe constructs
that differentiate between elements - for example,  prompting the user to create a construct that
differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to. The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For socialising' in order to capture the core distinction between the two system. The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
'For Science', and $5$ one that mainly serves  'socializing' purposes. Elements can also be rated with values between the poles if they may correspond to both or neither. The repertory grid software then identifies elements that are the least differentiated when requesting new constructs, in an attempt to elicit as much knowledge as possible from the user. We applied this technique to create an initial set of examples of social machines and to classify them according to specific features. To do so, we asked then computer science researchers familiar with the field to create their own repertory grids, and populate the elements from their own knowledge, and created the constructs using the standard repertory grid elicitation technique described above. This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
and $117$ different constructs (classifying factors). As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.

@@

\subsection{The polyarchical relationship of social machines}
When defining the boundaries of what we call social machines, one important observation was related to the distinction between platforms and technologies such as wikis and GWAP ('games-with-a-purpose'), which enable social machines to be created, and their instantiations into social machines that were brought to life for a specific purpose, such as Wikipedia and EteRNA\footnote{\url{http://eterna.cmu.edu}}, a game in which  participants design RNA  that is evaluated automatically using simulators. A second observation is concerned with the relationship between broader and narrower-scoped instances of social machines; general-purpose examples such as Facebook, Twitter or Amazon, or even the Web, enable the formation of more specific social machines and communities within them, that are working to solve solve different problems. An example illustrating this relationship is the Obama'12 US Presidential campaign~\cite{obamakieron}, which relied on Twitter and Facebook as basic social machines; most forms of customer relationship management and community engagement carried out in the digital space use similar tools. Finally, social machines are interconnected into a greater ecosystem, both at the social and at the computation levels. For example, insights on the classification of space objects gained through GalaxyZoo may influence the content of the corresponding articles, and the associated editorial processes, within Wikipedia. In the same time,a large number of communities within Wikipedia focus specific topics and activities, such as those contributing to science and technology.\footnote{Visualization of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}

Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behavior at various levels of granularity.  We propose looking at nested machines with 'The Web' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi) that spawn more specific social machines. A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as infrastructure, frameworks, services, and projects and initiatives.

@@

%One such example is the 'open access' movement in scholarly publishing
%~\cite{harnad2001self}. This cause mobilizes academic authors to self-archive their own papers online so they are available for free, in addition to hosted versions behind publishers pay-walls. In order to achieve this aim, the social machine of 'open access' typically uses the social machines of institutional repositories provided
%by the authors' associated institutions. As with other social machines, these social
%machines typically use off-the-shelf implementations (in this case the most popular are
%EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
%web. This cause is mature, growing year-on-year and has global coverage.

@@

\subsubsection{Tasks, purpose, and context of participation}
The first set of constructs pertain to \emph{what} the participants
do, both individually within the system, and collectively as a whole.
Individual activities might include creation of content, subjective
appraisal of existing content, posing of problems, solving of
problems, and so on.  The resulting 'computation', looked at from a
macroscopic whole, might be the identification of high-quality
artistic content, insight, or creative works, collective problem
solving, and so on.  Such tasks might be pre-defined by the system
designers, or brought to the system by participants.  Further, these
tasks might be of a specific type or class, or of many varied types or
encompass a general class of activities.  To capture this diversity,
we include a construct that differentiates machines with pre-defined problem spaces or are open to participant-provided tasks and activities.  Finally, we ask whether the benefit derived by
participants is distinct from that gained by machine/platform owners
and service providers.

@@

%% {\bf The types of tasks undertaken by the social component and the resulting contributions:}Social machines, just as any computing system, are built with a given purpose in mind, and exhibit functionality by which this purpose can be achieved. Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the 'computation'. This includes the creation and refinement of digital artifacts by executing given actions, possibly in collaboration with other participants. For example, in GalaxyZoo\footnote{\url{http://www.galaxyzoo.org}} people classify galaxies with collective performance as good as professional astronomers, while in reCaptcha\footnote{\url{http://www.google.com/recaptcha}} they refine the data that describes digital manuscripts. Sometimes, systems look for innovative ways of getting people to engage with the systems, including game narratives, like in EteRNA\footnote{\url{http://eterna.cmu.edu}}, in which  participants design RNA  that is evaluated automatically using simulators and then refined via human effort, or monetary rewards, like in Amazon's Mechanical Turk, a platform in which users undertake routine (micro) tasks such as object classification, audio transcription and text translation.

@@

From the dendrograms we can get a sense of similarities and separation of social machines. For
instance, it is clear that YouTube, Vimeo, Reddit, and Digg are similar, which is not unexpected, due to their focus on sharing content. Likewise, the exercise confirms commonalities between Quora and StackOverflow, which can be easily understood, given their question answering purpose. There is also a dendrogram for the constructs, which can be used, as more data is collected, as an indicator for correlation between constructs. In this particular case, dendrogram suggests that there is a correlation between the 'Single Participant Roles' and 'No Hierarchy in Participant Roles' constructs, which is because they refer to the same aspect of social machines. Another, but less obvious connection is between the 'Large variety of social features' and 'Benefit NOT for the world/society', which requires further investigation.

@@

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional

>>>>

social machines, overview of what they are, their history
    people interacting with content and with each other

@@

\section{Social machines: an introduction}
Once upon a time `machines' were programmed by programmers and used by users. The success of the Web has changed this relationship: we now see configurations of people interacting with content and with each other, typified by social Web sites. Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them  and  treat  each  such  compound  as  a  `social  machine'  -  a  machine in which the two aspects are seamlessly interwoven. This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.

Essentially social machines can be characterised as assemblies of manually executed and machine-driven (as in `automatised') services and the interaction of such services. A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship. When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines  because the users  are  part of the `social  computation', in this case the data creation and collection process facilitate through the site.  This social component  becomes  richer when  the database  is  curated by members  of  the broader community (e.g., Wikipedia) and when  the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through their individual or joint activities.

@@

This paper makes two main contributions: First, we introduce the notion of `social machines' and position it in connection to similar ideas and initiatives. Second we propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.

The remaining sections are organized as follows. In order to pin down what makes up a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} gives a summary of this comparative analysis. In Section \ref{sec:framework} we then present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via grid analysis. We conclude with an outlook of our future development activities, including details about the evaluation of the framework. In addition, we plan to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines who will contribute to the `computation' of a joint and widely agreed classification framework for the field.

\section{Comparison with related fields}
\label{sec:comparison}
Since the invention of the Turing machine we have experienced a `paradigm shift' in the usage of computers, moving away from purely calculative devices to facilitators of a wide range of human interactions. The emergence of `Computer-supported Collaboration Work (CSCW)' \cite{grudin1994computer} is representative for the early days of this trend. CSCW addresses `how collaborative activities and their coordination can be supported by means of computer systems' \cite{grudin1994computer}, whereas the initial concept evolved towards the more broader field of `Computer-supported collaboration' (CSC).

The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web. These developments led to amazing growth in terms of the amounts of content available online and mass participation. They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopaedias, publishing Terabytes of multimedia content, contributing to world-class software, and lively taking part in defining the agenda of many aspects of our society. This trend towards `prosumerism' is finding more and more adopters in the public and private sector as well, as governments and enterprises not only become active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organisational management, product development, services offers, and policies. In this context, a number of terms are used to refer to the ways people interact with each other and with applications: `wisdom of the crowds', `collective intelligence', `open innovation', `crowdsourcing', `human computation', and `social computing'. They are related, but not synonymous with `social machines', as explained in the following (see also Figure \ref{fig:socialmachine}).

@@

Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales hardly conceivable in the past, and to involve highly diverse and geographically distributed participants. A similar concept, though broader scoped, is collective intelligence, defined in  \cite{malone2009harnessing} as `groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as `social machines' is hidden in the the automation part of the latter. In a social machine, human and computational intelligence coalesce in order to achieve a given purpose. The overall setting may slightly vary in terms of the interaction patterns between the two types of services and the way their outcomes are used; however, where wisdom of the crowds and collective intelligence clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realization of hybrid systems where the two types of components co-exist. As such, theoretical and empirical insights from the two related areas are useful to understand the dynamics of the social structures underlying such a system, but they are definitely not the only ingredient needed to build operational social machines. The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering exercise by which a system is developed, tested, and updated are equally important.

One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a `a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}. At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, while in the same time ensuring in-time access of a potentially infinite pool of skills and resources not available before the advent of social computing technologies. The term crowdsourcing is typically associated with this larger collection of situations, in which `a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}. Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding. This sort of tasks are an important part of today's crowdsourcing projects landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work in the range of seconds to minutes to complete.  The key difference between social machines and open innovation and crowdsourcing is again in the computational part. The novelty these two approaches bring in lies in using a much larger pool of human resources as traditional work environments \cite{quinn2011human}. There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution. A principled combination of human and computational capabilities is less in the focus of such systems and the technical means that support them, by contrast to social machines. In comparison to human computation, social machines cover a wider range of scenarios. Human computation is AI-centric and uses people to perform tasks that computers are not (yet) able to tackle (in terms of accuracy) \cite{quinn2011human}; by comparison, we see many successful examples of social machines in which the role of machines is rather to facilitate interactions within groups of people or communities of interest.

An analogous line of reasoning could be followed to point out the overlaps between social machines and social computing \cite{parameswaran2007research}. The latter is an area of computer science which refers systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets' \cite{parameswaran2007research}. As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort. This distinction is even stronger in the case of social machines, which look at the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.

@@

\subsection{Methodology}
\label{sec:methodology}
The classification framework presented in this paper was created through knowledge elicitation ~\cite{knowledgeelicitation}. In particular we used the {\it repertory grid} elicitation
technique~\cite{kelly} in order to derive an initial set of {\it elements}, which represent instances of social machines, and {\it constructs}, which capture their most important characteristics.
In repertory grid elicitation a software tool is used to ask users to describe constructs
that differentiate between elements - for example,  prompting the user to create a construct that
differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to. The user describes the opposing poles of the construct - in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two system. The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
`For Science', and $5$ one that mainly serves  `socialising' purposes. Elements can also be rated with values between the poles if they may correspond to both or neither. The repertory grid software then identifies elements that are the least differentiated when requesting new constructs, in an attempt to elicit as much knowledge as possible from the user. We applied this technique to create an initial set of examples of social machines and to classify them according to specific features. To do so, we asked then computer science researchers familiar with the field to create their own repertory grids, and populate the elements from their own knowledge, and created the constructs using the standard repertory grid elicitation technique described above. This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
and $117$ different constructs (classifying factors). As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.

@@

\subsection{The polyarchical relationship of social machines}
When defining the boundaries of what we call social machines, one important observation was related to the distinction between platforms and technologies such as wikis and GWAP (`games-with-a-purpose'), which enable social machines to be created, and their instantiations into social machines that were brought to life for a specific purpose, such as Wikipedia and EteRNA\footnote{\url{http://eterna.cmu.edu}}, a game in which  participants design RNA  that is evaluated automatically using simulators. A second observation is concerned with the relationship between broader and narrower-scoped instances of social machines; general-purpose examples such as Facebook, Twitter or Amazon, or even the Web, enable the formation of more specific social machines and communities within them, that are working to solve solve different problems. An example illustrating this relationship is the Obama'12 US Presidential campaign~\cite{obamakieron}, which relied on Twitter and Facebook as basic social machines; most forms of customer relationship management and community engagement carried out in the digital space use similar tools. Finally, social machines are interconnected into a greater ecosystem, both at the social and at the computation levels. For example, insights on the classification of space objects gained through GalaxyZoo may influence the content of the corresponding articles, and the associated editorial processes, within Wikipedia. In the same time,a large number of communities within Wikipedia focus specific topics and activities, such as those contributing to science and technology.\footnote{Visualization of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}

Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behavior at various levels of granularity.  We propose looking at nested machines with `The Web' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi) that spawn more specific social machines. A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as infrastructure, frameworks, services, and projects and initiatives.

@@

%One such example is the `open access' movement in scholarly publishing
%~\cite{harnad2001self}. This cause mobilizes academic authors to self-archive their own papers online so they are available for free, in addition to hosted versions behind publishers pay-walls. In order to achieve this aim, the social machine of `open access' typically uses the social machines of institutional repositories provided
%by the authors' associated institutions. As with other social machines, these social
%machines typically use off-the-shelf implementations (in this case the most popular are
%EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
%web. This cause is mature, growing year-on-year and has global coverage.

@@

\subsubsection{Tasks, purpose, and context of participation}
The first set of constructs pertain to \emph{what} the participants
do, both individually within the system, and collectively as a whole.
Individual activities might include creation of content, subjective
appraisal of existing content, posing of problems, solving of
problems, and so on.  The resulting `computation', looked at from a
macroscopic whole, might be the identification of high-quality
artistic content, insight, or creative works, collective problem
solving, and so on.  Such tasks might be pre-defined by the system
designers, or brought to the system by participants.  Further, these
tasks might be of a specific type or class, or of many varied types or
encompass a general class of activities.  To capture this diversity,
we include a construct that differentiates machines with pre-defined problem spaces or are open to participant-provided tasks and activities.  Finally, we ask whether the benefit derived by
participants is distinct from that gained by machine/platform owners
and service providers.

@@

%% {\bf The types of tasks undertaken by the social component and the resulting contributions:}Social machines, just as any computing system, are built with a given purpose in mind, and exhibit functionality by which this purpose can be achieved. Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the `computation'. This includes the creation and refinement of digital artifacts by executing given actions, possibly in collaboration with other participants. For example, in GalaxyZoo\footnote{\url{http://www.galaxyzoo.org}} people classify galaxies with collective performance as good as professional astronomers, while in reCaptcha\footnote{\url{http://www.google.com/recaptcha}} they refine the data that describes digital manuscripts. Sometimes, systems look for innovative ways of getting people to engage with the systems, including game narratives, like in EteRNA\footnote{\url{http://eterna.cmu.edu}}, in which  participants design RNA  that is evaluated automatically using simulators and then refined via human effort, or monetary rewards, like in Amazon's Mechanical Turk, a platform in which users undertake routine (micro) tasks such as object classification, audio transcription and text translation.

@@

From the dendrograms we can get a sense of similarities and separation of social machines. For
instance, it is clear that YouTube, Vimeo, Reddit, and Digg are similar, which is not unexpected, due to their focus on sharing content. Likewise, the exercise confirms commonalities between Quora and StackOverflow, which can be easily understood, given their question answering purpose. There is also a dendrogram for the constructs, which can be used, as more data is collected, as an indicator for correlation between constructs. In this particular case, dendrogram suggests that there is a correlation between the `Single Participant Roles' and `No Hierarchy in Participant Roles' constructs, which is because they refer to the same aspect of social machines. Another, but less obvious connection is between the `Large variety of social features' and `Benefit NOT for the world/society', which requires further investigation.

@@

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs `a single self-contained file'!
%
%APPENDICES are optional
